---
title: "Methods in the Madness"
author: "Aaron Robotham"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Methods in the Madness}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

**Rfits** comes with a number of useful methods that operate on Rfits_image and Rfits_pointer objects. Using these you can do a lot of useful thing!

```{r}
library(Rfits)
library(Rwcs)
```

## Subsetting

First we will load up an image in RAM (temp_image) and as a pointer to the disk object (temp_point)

```{r}
file_image = system.file('extdata', 'image.fits', package = "Rfits")

temp_image = Rfits_read_image(file_image)
temp_point = Rfits_point(file_image)
```

We can look at both of these, since they have distinct print methods by default:

```{r}
print(temp_image)
print(temp_point)
```

The two data structures are a bit different, but we can still plot both directly:

```{r fig.width=6, fig.height=6}
plot(temp_image)
plot(temp_point)
```

Behind the scene the S3 plot method does different things depending on the Rfits object encountered. If it is an Rfits_image then it will directly plot it using *Rwcs_image* with a WCS. If it is a Rfits_pointer then it will first fully load the data into RAM, and then plot it using *Rwcs_image*.

If you have a Rfits_pointer object then you can convert it to a Rfits_image very easily:

```{r}
temp_point[]
```

This is because Rfits has special subset operation that work specifically on Rfits_image and Rfits_pointer objects. This means we can plot a subset:

```{r fig.width=6, fig.height=6}
plot(temp_image[50:150,50:150])
plot(temp_point[50:150,50:150])
```

For convenience the following will achieve the same outcomes:

```{r fig.width=6, fig.height=6}
plot(temp_image[c(50,150),c(50,150)])
plot(temp_point[c(50,150),c(50,150)])
```

And we can also cutout a box on a specific location:

```{r fig.width=6, fig.height=6}
plot(temp_image[101, 101, box=100])
plot(temp_point[101, 101, box=100])
```

We can also subset based on RA/Dec coordinates:

```{r}
coords = Rwcs_p2s(101,101,temp_image$keyvalues)
print(coords)
```

```{r fig.width=6, fig.height=6}
plot(temp_image[coords[1], coords[2], box=100, type='coord'])
plot(temp_point[coords[1], coords[2], box=100, type='coord'])
```

## Useful Class Methods

Rfits comes with a whole lot of specific Rfits_image and Rfits_pointer class methods (for details see ?Rfits_methods):

*   print
*   length (total length of the image/cube/array)
*   dim (dimension of the image/cube/array)
*   centre/center (RA, Dec of image centre)
*   corners (RA, Dec of image corners)
*   extremes (RA, Dec of image extremes)
*   pixscale (pixel scale of centre of image)
*   pixarea (pixel area of centre of image)
*   rotation (rotation of image in degrees)

They are designed to give identical results, i.e.:

```{r}
print(temp_image)
print(temp_point)
```

```{r}
length(temp_image)
length(temp_point)
```

```{r}
dim(temp_image)
dim(temp_point)
```

```{r}
centre(temp_image)
centre(temp_point)
```

```{r}
corners(temp_image)
corners(temp_point)
```

```{r}
extremes(temp_image)
extremes(temp_point)
```

```{r}
pixscale(temp_image)
pixscale(temp_point)
```

```{r}
pixarea(temp_image)
pixarea(temp_point)
```

```{r}
rotation(temp_image)
rotation(temp_point)
```

## Pixel Manipulation Methods

Rfits_image objects can be usefully operated on as if they were matrices:

```{r fig.width=6, fig.height=6}
plot(temp_image^5, qdiff=TRUE)
plot(temp_image - 10, qdiff=TRUE)
```

Most **R** mathematical operators will work. You can also directly operate with Rfits_image and Rfits_point objects:

```{r fig.width=6, fig.height=6}
plot(temp_image * temp_image, qdiff=TRUE)
plot(temp_image - temp_image, qdiff=TRUE) #blank, since the image should now be 0
plot(temp_point - temp_point, qdiff=TRUE) #blank, since the image should now be 0
```

If you mix Rfits object types you will need to convert the Rfits_pointer to a Rfits_image using the subset operation:

```{r fig.width=6, fig.height=6}
plot(temp_image * temp_point[], qdiff=TRUE)
plot(temp_image - temp_point[], qdiff=TRUE) #blank, since the image should now be 0
```
